Practical Challenges with Web Service Evolution:\\A Review of Literature and Current Industry Practices

### 2013-06-05 paper.tex
almost there, 15 page with the modifications

However, in todays technology multiple messaging formats in addition to XML are used. Wikipedia provides a generic definition of web services as follows: \cite{web_service_wiki}:

@@

For example, a geographical mapping web service may be consumed by many different kinds of consumers in real life such as: real-estate listing web sites, local business search portals, transit information web sites, etc. Similarly, payment processing web services are used by a multitude of consumers such as: e-commerce applications, retail stations etc.

To carry out the actual communication, the web service publishers and consumers follow specific protocols. Of the many protocols that are in use today \cite{web_service_protocols_wiki}, two commonly used protocols are as follows:

@@

\subsection{SOAP Web Services} % (fold)
\label{sub:soap_web_services}
SOAP was designed in 1998 by Winer et al. for Microsoft and is currently maintained by the World Wide Web Consortium \cite{soap_wiki}. In brief, SOAP is an XML based protocol to communicate structured data and necessary headers between the consumers and publisher of a web service.

A SOAP web service is composed of one or more operations or methods. SOAP XML messages are used for communication between a consumer of the web service and its methods. The structure of these methods, their corresponding messages and other meta information of a SOAP service can be described using an accompanying XML based Web Services Description Language or WSDL file. In addition to providing a structured definition of the service, these WSDL files can be automatically parsed by code-generators to provide the consumers with an easy to integrate interface.

@@

\subsection{RESTful Web Services} % (fold)
\label{sub:restful_web_services}
Being an XML based protocol, SOAP messages are often verbose. Also, the additional header information introduces communication overheads. On the other hand, RESTful web services gained popularity with the rise of cloud computing as they provide a lightweight and flexible approach \cite{mangler2010origin}, where the HTTP protocol stack is leveraged without imposing constraints on the actual format of the messaging.

The term Representational State Transfer (REST) was introduced by Roy Fielding \cite{rest_wiki} and is used interchangeably as RESTful web service or Web API. In brief, REST is an implementation for web services where user defined ``resources'' are created, accessed, modified and deleted utilizing the various features of the HTTP stack such as caching, security, layering and different request methods (GET, POST, PUT, DELETE etc.), content types.

@@

The scope of this paper is limited to only the evolutionary aspects of web services. On this topic, the remainder of this paper is organized as follows: in the next section we discuss the motivation behind this research followed by our research questions. Then, we discuss the research method and findings in Sections \ref{sec:research_method} and \ref{sec:research_findings} respectively. Finally, we present the key challenges and future work opportunities with evolving web services in Section \ref{sec:discussion}.

\section{Motivation} % (fold)
\label{sec:motivation}
Web services, like other software systems, need to evolve to stay on top of bugs and ever changing requirements. For example, between May 15th 2013 and May 22nd 2013, the Facebook API team saw 238 new bugs, resolved 211 bugs and fixed 24 bugs \cite{facebook_release_note}. However, often times web services are used by an unknown number of consumers, and the web service publishers have little control over them. Updating a web service can be tricky in such a situation since a breaking change may upset the existing consumers. To be able to evolve in such a decoupled scenario, multiple versions of a service need to coexist which poses both technical and business challenges in multiple fronts. At a conceptual level, Figure \ref{fig:web_service_layers} shows the different parts of a web service that need to deal with evolution challenges:

@@

Focusing on these questions helped us identify some of the gaps of the currently available techniques for evolving web services. Using this as a background, innovative future work can be carried out to fill these gaps and simplify web service evolution.
% section research_questions (end)

@@

We have performed a Systematic Literature Review to extract information from the published articles following \cite{kitchenham2007guidelines}. We have also investigated three published real-word web services to find out the answers to the research questions based on todays solutions.
In this regard, we have limited our research to two high level sources as follows:

@@

From the search results, we performed a primary screening and selected 21 papers after reading the abstracts. Then, for the remaining articles, we read the full paper as long as it was found relevant, i.e. only if it contained topics about the evolution of web services. The resulting list is composed of \cite{kitchenham2007guidelines} \cite{petersen2008systematic} \cite{treiber2009analyzing} \cite{borovskiy2008evolution} \cite{lublinsky2007versioning} \cite{laskey2008considerations} \cite{wilde2004semantically} \cite{fang2007version} \cite{leitner2008end} \cite{mangler2010origin} \cite{kaminski2006design} \cite{le2008synchronizing} \cite{aversano2005visualizing}.

@@

One or more of these roles may be played by one individual. The stakeholders take part in driving changes to a web service. The authors identified the following types of changes in a web services: Interface, Implementation, Quality of Service (QoS), Usage, Requirement, Service-level Agreement (SLA), Pre and post conditions, and Feedback. In line with \cite{treiber2009analyzing}  Lublinsky also pointed out that a new version may be required primarily because of changes in its interface, schema of the message payload or in it’s implementation. In \cite{laskey2008considerations}, Laskey identified the key reasons for changes between versions as: a) service description changes, b) changes in functions, c) changes in the interaction mechanics, and d) change of known pre and post conditions. These changes would require new versions of the service as well as accompanying any explanation of the specific changes.

In addition to the types of changes discussed above, Borovskiy et al. identified some drivers behind web service evolution from the perspective of an enterprise, where both the web service consumer and publisher may be part of the same organization \cite{borovskiy2008evolution}. In this regard, they categorized two classes of drivers behind the evolution of web services as follows:

@@

A change in a web service impacts one or more of the stakeholder roles based on its type. Treiber et al. pointed out that, in addition to stakeholder impacts, the sources are also interdependent on each other \cite{treiber2009analyzing}. For example, when a web service \textbf{provider} decides to change the \textbf{interface}, the \textbf{developer} needs to write the changed interface and its \textbf{implementation}. If the changed interface if not backward compatible, and the service \textbf{integrator} needs to modify the integration. The new interface and implementation may instigate a change its \textbf{QoS}, since it may improve or degrade the performance. The changes may improve or degrade the \textbf{QoS} and impact the \textbf{end user}.

@@

  \item \textbf{Units of versioning}: One approach to version a service is to apply a single version to all it’s methods that can be deployed as a group. Another approach is to version individual methods or operations of the service as they change, this confines the change to only specific methods and provides a finer grained versioning. However, deployment of such versioned methods are complicated and it forces the consumer to specify the service, its method name and an additional version to use for each operation.
  \item \textbf{Service version life-cycle considerations}: Based on the actual need, Lublinsky suggests adjusting the life-cycle of versions, as too many versions increase the maintenance overhead while too few versions leave the consumers a smaller time window to upgrade since each version has a shorter life-cycle.
  \item \textbf{Version deployment/access approaches}: When deploying versioned services, Lublinsky compares between multiple versions deployed on the same endpoint vs. each version deployed on its own endpoint. Lublinsky points out that the former approach may introduce conflicts in names of the objects, databases and other related resources, and requires intermediate routers, which typically lowers performance and may introduce SLA complications. The author favored the latter approach and argued that it provides better scalability and loose coupling between versions.

@@

To summarize, this paper paper points out some key trade-offs that need to be addressed while evolving web services based on the specific use case.

A versioning scheme is required to identify and use a specific version. Laskey discusses about the versioning of web services in \cite{laskey2008considerations}. In this paper, the author discussed versions in terms of identifiable resources, that can have multiple versions and are modified by applying new versions to the original one.

To identify a specific version of a resource, Leskey suggests using a unique identifier with an accompanying explanation of how to interpret it. The versioning identifiers should be used consistently. In addition to this, the author also recommends specifying a policy describing compatibility between the versions so the consumers clearly understand the impacts of version changes. As an example, the author mentioned a generic policy as follows:

@@

@@

\subsubsection{Protocol Level Implementation:}
Since web services typically follow a protocols such as SOAP, REST, etc., protocol level support for evolving web services would allow for uniform standards. In \cite{wilde2004semantically} Wilde proposed a conceptual framework for SOAP web services where consumers can be forward compatible in a managed way.

At it’s core, the framework includes open and extensible XML schemas so the consumers of old version can work with the schema in a newer version. In addition to this, the framework recommends using a set of custom extensions to the XML schema. These extensions can be as simple as XML tags such as mustUnderstand, mayIgnore or some complex ones to suit the specific use case in hand. But irrespective of their actual usage, an accompanying guideline on their processing rules need to be provided. Using the approach of this framework, the consumers of the web service can attain a predictable level of ``graceful degradation'' as the service evolves.

Using a similar approach but instead of arbitrary custom tags to extend the schema, Fang et al. provides a concrete list of six XML nodes to create a version-aware service model for SOAP web services \cite{fang2007version}. They introduced extensions to WSDL and UDDI to tackle evolving web services. These extensions are composed of six XML nodes: \textbf{1) version name, 2 ) version description, 3) startTime of the version, 4) endTime of the version, 5) alias such as (new, current etc) of the version name and, 6) original version that preceded it}.

With the help of this versioning description of the service, a consumer can find and register for a target version through a service registry. The server can publish events as new a version of a service is deployed and the registered consumers can subscribe to these events. When a new version is released, the consumer gets notified about the change and can self update if the changes are backward compatible. The authors discussed an example implementation of this approach and demonstrated that using this approach SOAP services can be versioned and future changes can be communicated to the consumers.

@@

\begin{figure*}[ht]
  \centering
  \includegraphics[width=\textwidth]{vresco.png}
    \caption{VRESCo Service Version Graph, source: \cite{leitner2008end}}
  \label{fig:vresco}
\end{figure*}

For each version, VRESCo uses a free-form tag to identify it. The version tag and associated meta information from the WSDL file of each version is stored in VRESCo's database. So, when new versions are released, it can perform an analysis to see if the changes are backward compatible. To route a consumer request to the right version, the versions of a service are deployed behind a proxy. Based on a ``selection strategy'' specified by the consumer (e.g. always use latest, always use stable, fix version, etc.) the proxy can then automatically select the desired version without needing the consumer to change it’s implementation. For compatible updates, this approach off-loads the consumer at the expense of an additional proxy layer.

@@

@@

Following from the protocol level implementation approaches discussed in the previous section, we have presented the application code level solution approaches in this section.

@@

\begin{figure*}[ht]
  \centering
  \includegraphics[width=\textwidth]{chain_of_adapters.png}
    \caption{Chain of Adapters structure, source: \cite{kaminski2006design}}
  \label{fig:chain_of_adapters}
\end{figure*}

@@

The ``Chain of Adapters'' technique addresses their requirements, but they mentioned some caveats. For example, the adapters need to ensure all breaking changes are compensated, which my need for a separate data store when fields are removed from the datastore on subsequent versions. Also, this chain effectively connects all the versions, so a bug or it’s fix on one version may have impact on all the versions.

@@

In \cite{le2008synchronizing} Zou et al. presents a technique to generate consumer specific changelogs for evolving SOAP web services. For existing consumers, the authors presented an approach where a consumer specific changelog can be generated based on their usage pattern of the service. The approach is described below:

\begin{itemize}
  \item \textbf{Service invocation monitoring}: a monitoring system is used to track the usage pattern of each consumer so when a service is changed, it can identify specific areas of interest for a consumer
  \item \textbf{Service difference analysis}: Since SOAP services are described using XML based WSDL files, the differences between the two versions such as addition, removal or modification can be analyzed by comparing the WSDL files
  \item \textbf{Release note customization}: The release note can then be tailored for each consumer by using the data collected at step 1 to only include relevant changes per customer
  \item \textbf{Code release note linkage}: Using the customized release note from step 3, it is possible automatically inspect the consumer code and discover fragments of the code that need to change for the new version.
\end{itemize}

@@

@@

\begin{itemize}
  \item \textbf{Google Maps}: Google Maps provides a web service for its consumers and updates the service to provide ``new features, bug fixes and performance improvements'' \cite{google_maps_versioning}.
  \item \textbf{Twitter Web API}: Twitter provides a RESTful web service to interact with its data from third-party applications \cite{twitter_api}. This web service is also evolving to fix bugs and bring new features, but using a different approach from that of Google Maps.
  \item \textbf{Facebook Graph API}: Facebook Graph API is the primary web service to retrieve and post data to Facebook \cite{facebook_api}. The service evolves to support change/removal of functionality, backwards compatibility, Facebook product changes, and privacy and security related changes.
\end{itemize}

@@

\begin{itemize}
  \item \textbf{Experimental version}: With all the latest features, bug fixes and improvements. But this version is not guaranteed to be feature stable.
  \item \textbf{Release version}: From the the experimental version, every quarter, a feature stable version of the service is tagged as a ``release'' version. These release versions may receive bug fixes but the features remain stable.
  \item \textbf{Previously released or Frozen version}: Every time a new ``release version'' is tagged, the oldest existing release version is ``frozen'', so no new features or bug fixes are applied to ensure it remains completely feature stable.
\end{itemize}
Every time a new release version is tagged, the oldest version is retired. Google Maps guarantees the versions are backward compatible and automatically migrates the consumers using the oldest version to the following version without needing the consumers to make any change on their ends.

On the other hand, the Twitter API had two versions at the time of writing this paper \cite{twitter_api} and the old version is marked as deprecated. The users of the old version are suggested to upgrade to the new version before its posted date of expiry. Once a version is released, future changes are made to the released version to provide new features and bug fixes. However, unlike Google Maps, new features are not always released under a new version name. Instead, the changes may be deployed under the same version name.

Unlike the previous approaches, Facebook Graph API on the other hand uses a different approach for its evolution. They have a 90-day breaking change policy stated as follows: \cite{facebook_api}

@@

@@

\subsubsection{Documentation:}
Google Maps web service provides a human readable documentation of the service operations with example requests and responses specific to a version \cite{google_maps_services}. Since the API is publicly available, the users can run the example requests to produce live results by browsing the API endpoints on their browsers.

Twitter also publishes human readable API documentation for each version of the web service in their developer portal \cite{twitter_api}. The portal also includes an interactive web browser based console to let its users explore the service on the browser \cite{twitter_console}.

@@

@@

@@

Facebook also uses a developer portal, a StackOverflow channel and blogs to communicate with its users \cite{facebook_stack_overflow}. Among other things, the developer portal announces upcoming changes and effective dates of the 90-day breaking changes.
% subsection review_of_industry_examples (end)

@@

Based on our findings as discussed in the preceding sections, in this section we have summarized the key challenges and identified avenues for future work related to evolving web services. The challenges are as follows:

\subsection{Deployment Challenges} % (fold)
\label{sub:deployment_challenges}

Multiple versions of an evolving web service need to be deployed at the same time to support the consumers of each version. This poses the following challenges:
\begin{itemize}
  \item Should multiple versions be deployed under a separate or a single endpoint?
  \item Should a version apply to a whole service or to the individual methods?
  \item How many versions need to be deployed at any given time?
  \item How and when to migrate the consumers to a newer version?
  \item How and when to migrate the consumers that are using a retiring version?
  \item How to identify and interpret a version?
\end{itemize}

% subsection deployment_challenges (end)

\subsection{Implementation Challenges} % (fold)
\label{sub:implementation_challenges}

The implementation of a web service needs to be suitable for it’s evolution. In this regard, we have identified the following challenges:
\begin{itemize}
  \item How to design and organize the code for evolving web services?
  \item How to provide protocol level support for versioning?
  \item How to fix defects that affect multiple versions?
  \item How to design the data storage for evolving web services?
  \item How to automatically test multiple versions of a web service?
  \item How to implement sandbox environment for testing different versions?
\end{itemize}

% subsection implementation_challenges (end)

\subsection{Documentation Challenges} % (fold)
\label{ssub:documentation_challenges}
A documentation of the web service is used by the consumer to understand and explore the features of the web service. Evolving web services introduce several challenges related to their documentation as follows:
\begin{itemize}
  \item How to produce version specific documentation?
  \item How to auto generate human readable usage examples of a web service?
  \item How to provide an interactive explorer for different versions a web service?
  \item How to generate changelogs between versions?
\end{itemize}
% subsubsection documentation_challenges (end)

\subsection{Communication Challenges} % (fold)
\label{sub:communication_challenges}
Communication plays a big role behind an evolving web service because the communication channels are used to share information about future changes as well as to collect feedback about defects and feature requests. The following is a list of communication related challenges with an evolving web service:
\begin{itemize}
  \item What communication channels should be used?
  \item How to automatically notify the consumers about the changes?
  \item If multiple channels are used, how to aggregate the data from multiple channels?
\end{itemize}

% subsection communication_challenges (end)

\subsection{Future Work} % (fold)
\label{sub:future_work}
% subsection future_work (end)
From our review of the existing literature and industry practices, we have already discussed their suggested approaches to solve some of the aforementioned challenges. Even though these challenges are generic in nature, we recognize that the answers to these challenges may rely on the specific use-case of a web service. However, due to the lack of standard approaches and tool support to address these challenges, once a decision is made, it typically requires custom implementation to solve these general problems. We identify this lack of standards and tool support as an opportunity for innovative future work to simplify the evolution of web services.
% section discussion (end)

@@

In this paper, from our systematic literature reviews, we have identified the change drivers behind evolving web services and several approaches to deal with the various aspects of an evolving web service, such as deployment, versioning, protocol and source code level implementation, documentation and communication. In addition to the literature review, we have found different approaches to deal with these challenges from three real world popular web services.

Based on the review, we have summarized our findings of the key challenges with evolutionary web services in four high level categories a) Deployment, b) Implementation, c) Documentation and, d) Communication. For each of these categories, we have compiled a shortlist of challenges in question forms. This list of questions can be treated as a checklist for working on evolving web services. We have also discussed different approaches to solve some of the challenges from our review of the existing literature and industry practices. In our future work, we aim to reconcile and augment these fragmented solutions to provide a standard approach and tool support for evolving web services.
% section conclusion (end)

>>>>

However, in todays technology, a web service is not limited to XML format and apply to a more generic context as defined in Wikipedia: \cite{web_service_wiki}:

@@

For example, a geographical mapping web service may be consumed by many different kinds of consumers in real life such as: real-estate listing web sites, local business search portals, transit information web sites, etc.

To carry out the actual communication, a web service may follow a protocol. Of the many protocols that are in use today \cite{web_service_protocols_wiki}, two popular protocols are:

@@

\subsection{SOAP Web Services} % (fold)
\label{sub:soap_web_services}
SOAP was designed in 1998 by Winer et al. and is currently maintained by the World Wide Web Consortium \cite{soap_wiki}. In brief, SOAP is an XML based protocol to communicate structured data and necessary headers between the consumers and publisher of a web service.

A SOAP web service is composed of one or more operations or methods. SOAP XML messages are used for communication between a consumer of the web service and its methods. The structure a SOAP service and its messages can be described using an accompanying XML based Web Services Description Language (WSDL) file. In addition to providing a structured definition of the service, these WSDL files can be automatically parsed by code-generators to provide the consumers with an easy-to-integrate interface.

@@

\subsection{RESTful Web Services} % (fold)
\label{sub:restful_web_services}
Being an XML based protocol, SOAP messages are often verbose. On the other hand, RESTful web services provide a lightweight and flexible approach, and gained popularity with the rise of cloud computing \cite{mangler2010origin}.

The term Representational State Transfer (REST) was introduced by Roy Fielding \cite{rest_wiki} and is used interchangeably as RESTful web service or Web API. In brief, REST is an implementation for web services where user defined ``resources'' are created, accessed, modified and deleted utilizing the various features of the HTTP stack such as caching, security, layering and different request methods (GET, POST, PUT, DELETE etc.), and content types.

@@

The scope of this paper is limited to only the evolutionary aspects of web services. On this topic, the remainder of this paper is organized as follows: in the next section we discuss the motivation behind this research followed by our research questions. Then, we discuss the research method and findings in Sections \ref{sec:research_method} and \ref{sec:research_findings} respectively. Finally, we summarize the key challenges and future work opportunities with evolving web services in Section \ref{sec:discussion}.

\section{Motivation} % (fold)
\label{sec:motivation}
Web services, like other software systems, need to evolve to stay on top of bugs and ever changing requirements. For example, between May 15th 2013 and May 22nd 2013, the Facebook API team saw 238 new bugs, resolved 211 bugs and fixed 24 bugs \cite{facebook_release_note}. However, often times web services are used by an unknown number of consumers, and the web service publishers have little control over them. Updating a web service can be tricky in such a situation since a breaking change may upset the existing consumers. To be able to evolve in such a decoupled scenario, multiple versions of a service need to coexist which poses both technical and business challenges in multiple fronts. At a conceptual level, Figure \ref{fig:web_service_layers} shows the different parts of a web service that need to deal with evolution:

@@

Focusing on these questions helped us identify gaps of the currently available techniques for evolving web services. Using this as a background, innovative future work can be carried out to fill these gaps and simplify web service evolution.
% section research_questions (end)

@@

We have performed a Systematic Literature Review to extract information from the published articles following \cite{kitchenham2007guidelines}. We have also analyzed three real-word web services to find out the answers to the research questions based on industry examples.
In this regard, we have limited our research to two high level sources as follows:

@@

From the search results, we performed a primary screening and selected 22 papers after reading the abstracts. Then, for the remaining articles, we read the full paper as long as it was found relevant, i.e. only if it contained topics about the evolution of web services. The resulting list is composed of \cite{kitchenham2007guidelines} \cite{petersen2008systematic} \cite{treiber2009analyzing} \cite{borovskiy2008evolution} \cite{lublinsky2007versioning} \cite{laskey2008considerations} \cite{wilde2004semantically} \cite{fang2007version} \cite{leitner2008end} \cite{mangler2010origin} \cite{kaminski2006design} \cite{le2008synchronizing} \cite{aversano2005visualizing}.

@@

The stakeholders take part in driving changes to a web service. The authors identified the following types of changes in a web services: Interface, Implementation, Quality of Service (QoS), Usage, Requirement, Service-level Agreement (SLA), Pre and post conditions, and Feedback. In line with \cite{treiber2009analyzing}  Lublinsky also pointed out that a new version may be required primarily because of changes in its interface, schema of the message payload or in it’s implementation. In \cite{laskey2008considerations}, Laskey identified the key reasons for changes between versions as: a) service description changes, b) changes in functions, c) changes in the interaction mechanics, and d) change of known pre and post conditions. These changes would require new versions of the service as well as accompanying any explanation of the specific changes.

In addition to the types of changes discussed above, Borovskiy et al. identified some drivers behind web service evolution from the perspective of an enterprise \cite{borovskiy2008evolution}. They categorized two classes of change drivers:

@@

A change on a web service impacts one or more of the stakeholders. Treiber et al. pointed out that, the change drivers are also interdependent on each other \cite{treiber2009analyzing}. For example, when a web service \textbf{provider} decides to change the \textbf{interface}, the \textbf{developer} needs to write the changed interface and its \textbf{implementation}. If the changed interface is not backward compatible, and the service \textbf{integrator} needs to modify the integration. The new interface and implementation may instigate a change its \textbf{QoS}, since it may improve or degrade the performance. The changes may improve or degrade the \textbf{QoS} and impact the \textbf{end user}.

@@

  \item \textbf{Units of versioning}: One approach to version a service is to apply a single version to all it’s methods that can be deployed as a group. Another approach is to version individual methods or operations of the service as they change, this confines the change to only specific methods and provides a finer grained versioning. However, deployment of such versioned methods are complicated and requires the consumer to specify a version for each method of the service.
  \item \textbf{Service version life-cycle considerations}: Lublinsky suggests adjusting the life-cycle of versions based on the actual need. Too many versions require more maintenance while too few versions leave the consumers a smaller time window to upgrade since each version has a shorter life-cycle.
  \item \textbf{Version deployment/access approaches}: When deploying versioned services, Lublinsky compares between multiple versions deployed on the same endpoint vs. each version deployed on its own endpoint. Lublinsky favors the latter approach as it doesn't need intermediate routers and provides better scalability and loose coupling between versions.

@@

A versioning scheme is required to identify and use a specific version. In \cite{laskey2008considerations} Laskey discussed versions in terms of identifiable resources, that can have multiple versions and are modified by applying new versions to the original one.

To identify a specific version of a resource, Leskey suggests using a unique identifier with an accompanying explanation of how to interpret it. The versioning identifiers should be used consistently. In addition to this, the author also recommends specifying a policy describing compatibility between the versions as shown in the following example:

@@

From these versioning related papers, we identify the following \textbf{Deployment Challenges} for evolving web services:

\begin{itemize}
  \item Should multiple versions be deployed under a separate or a single endpoint?
  \item Should a version apply to a whole service or to the individual methods?
  \item How many versions need to be deployed at any given time?
  \item How to identify and interpret a version?
\end{itemize}

@@

\subsubsection{Protocol Level Implementation:}
Since web services typically follow protocols such as SOAP, REST, etc., protocol level support for evolving web services would allow for uniform standards. In \cite{wilde2004semantically} Wilde proposed a conceptual framework for SOAP web services where consumers can be forward compatible in a managed way.

At it’s core, the framework includes open and extensible XML schemas so the consumers of old version can work with the schema in a newer version. In addition to this, the framework recommends using a set of custom extensions to the XML schema. These extensions can be as simple as XML tags such as mustUnderstand, mayIgnore or some complex ones to suit the specific use case in hand. To communicate their actual usage, an accompanying guideline on their processing rules need to be provided so that the consumers of the web service can attain a predictable level of ``graceful degradation'' as the service evolves.

Using a similar approach but instead of arbitrary custom tags to extend the schema, Fang et al. provides a concrete list of six XML nodes to create a version-aware service model for SOAP web services \cite{fang2007version}. They introduced extensions to WSDL and UDDI \cite{uddi_wiki} to tackle evolving web services. The six XML nodes are: \textbf{1) version name, 2 ) version description, 3) startTime of the version, 4) endTime of the version, 5) alias such as (new, current etc) of the version name and, 6) original version that preceded it}.

Using these tags, a consumer can find and register for a target version through a service registry. When a new version is released, a registered consumer gets notified about the change and can self update if the changes are backward compatible.

@@

\begin{figure*}[bt]
  \centering
  \includegraphics[scale=0.35]{vresco.png}
    \caption{VRESCo Service Version Graph, source: \cite{leitner2008end}}
  \label{fig:vresco}
\end{figure*}

For each version, VRESCo uses a free-form tag to identify it. The version tag and associated meta information from the WSDL file of each version is stored in VRESCo's database. So, when new versions are released, it can analyze to see if the changes are backward compatible. To route a consumer request to the right version, the versions of a service are deployed behind a proxy. Based on a ``selection strategy'' specified by the consumer (e.g. always use latest, always use stable, fix version, etc.) the proxy can then automatically select the desired version. For compatible updates, this approach off-loads the consumer at the expense of an additional proxy layer.

@@

From these different approaches, we have identified the following challenges with evolving web services:
\begin{itemize}
  \item \textbf{Implementation Challenge}: How to provide protocol level support for versioning?
  \item \textbf{Communication Challenge}: How to automatically notify the consumers about the changes?

@@

Following from the protocol level implementation approaches discussed in the previous section, we have presented the application level solution approaches in this section.

@@

\begin{figure*}[ht]
  \centering
  \includegraphics[scale=0.4]{chain_of_adapters.png}
    \caption{Chain of Adapters structure, source: \cite{kaminski2006design}}
  \label{fig:chain_of_adapters}
\end{figure*}

@@

The ``Chain of Adapters'' technique addresses their requirements, but they mentioned some caveats. For example, the adapters need to ensure all breaking changes are compensated, which my need for a separate data store when fields are removed from the datastore on subsequent versions. Also, this chain effectively connects all the versions, so a bug or its fix on one version may have impact on all the versions.

From this study, we have compiled the following list of \textbf{Implementation Challenges} for evolving web services:

\begin{itemize}
  \item How to design and organize the code for evolving web services?
  \item How to fix defects that affect multiple versions?
  \item How to design the data storage for evolving web services?
  \item How to automatically test multiple versions of a web service?
\end{itemize}

@@

In \cite{le2008synchronizing} Zou et al. presents a technique to generate consumer specific changelogs for evolving SOAP web services. The authors presented an approach where a consumer specific changelog can be generated based on their usage pattern of the service. The approach is described below:

\begin{itemize}
  \item \textbf{Service invocation monitoring}: A monitoring system tracks the usage pattern of each consumer to identify specific areas of interest for a consumer.
  \item \textbf{Service difference analysis}: The differences between two versions of a web service is analyzed by comparing their WSDL files.
  \item \textbf{Release note customization}: The release note can then be tailored for each consumer by using the data collected at step 1 to only include relevant changes per customer
  \item \textbf{Code release note linkage}: Using the customized release note from step 3, it is possible automatically inspect the consumer code and discover fragments of the code that need to change for the new version.
\end{itemize}

@@

From these papers, we compiled the following list of \textbf{Documentation Challenges} of evolving web services:

\begin{itemize}
  \item How to produce version specific documentation?
  \item How to generate changelog between versions?
\end{itemize}

@@

\begin{itemize}
  \item \textbf{Google Maps}: Google Maps provides a web service for its consumers and updates the service to provide ``new features, bug fixes and performance improvements'' \cite{google_maps_versioning}.
  \item \textbf{Twitter Web API}: Twitter provides a RESTful web service to interact with its data from third-party applications \cite{twitter_api}. This web service is also evolving to fix bugs and introduce new features.
  \item \textbf{Facebook Graph API}: Facebook Graph API is the primary web service to retrieve and post data to Facebook \cite{facebook_api}. The service evolves to support change/removal of functionality, backwards compatibility, Facebook product changes, and privacy and security related changes.
\end{itemize}

@@

\begin{itemize}
  \item \textbf{Experimental version}: With all the latest features, bug fixes and improvements. But this version is not guaranteed to be feature stable.
  \item \textbf{Release version}: From the the experimental version, every quarter, a feature stable version of the service is tagged as a ``release'' version. These release versions may receive bug fixes but the features remain stable.
  \item \textbf{Previously released or Frozen version}: Every time a new ``release version'' is tagged, the oldest existing release version is ``frozen'', so no new features or bug fixes are applied to ensure it remains completely feature stable.
\end{itemize}
Every time a new release version is tagged, the oldest version is retired. Google Maps guarantees the versions are backward compatible and automatically migrates the consumers using the oldest version to the following version.

On the other hand, the Twitter API had two versions at the time of writing this paper \cite{twitter_api} and the old version is marked as deprecated. Once a version is released, future changes are made to the released version to provide new features and bug fixes. However, unlike Google Maps, new features are not always released under a new version name. Instead, the changes may be deployed under the same version name.

Unlike the previous approaches, Facebook Graph API uses a 90-day breaking change policy as follows: \cite{facebook_api}

@@

From these different approaches, we compiled the following list of \textbf{deployment challenges} to evolving web services:

\begin{itemize}
  \item How and when to migrate the consumers to a newer version?
  \item How and when to migrate the consumers that are using a retiring version?
\end{itemize}

@@

\subsubsection{Documentation:}
Google Maps web service provides a human readable documentation of the service operations with examples for each version \cite{google_maps_services}. Since the API is publicly available, the users can run the example requests to produce live results by browsing the API endpoints on their browsers.

Twitter also publishes human readable API documentation for each version of the web service in their developer portal \cite{twitter_api}. The portal includes an interactive web browser based console to let its users explore the service on the browser \cite{twitter_console}.

@@

From these examples, we have compiled the following list of \textbf{Documentation Challenges} of evolving web services:

\begin{itemize}
  \item How to auto generate human readable usage examples of a web service?
  \item How to provide an interactive explorer for different versions a web service?
\end{itemize}

@@

Facebook also uses a developer portal, a StackOverflow channel and blogs to communicate with its users \cite{facebook_stack_overflow}.

@@

From these communication patterns, we identified the following list of \textbf{Communication Challenges}:

\begin{itemize}
  \item What communication channels should be used?
  \item If multiple channels are used, how to aggregate the data from multiple channels?
\end{itemize}

% subsection review_of_industry_examples (end)

@@

In the previous sections, we have provided a list of different challenges in question formats related to evolving web services in 4 categories:

\begin{itemize}
  \item Deployment Challenges
  \item Implementation Challenges
  \item Documentation Challenges
  \item Communication Challenges
\end{itemize}

These challenges are derived from our systematic review of the literature and analysis of three real world web services. We have also listed suggested approaches to tackle some of the problems. Even though the answers to these challenges rely on specific use-case of a web service, we recognize that these challenges are generic in nature. However, due to the lack of standard approaches to address these challenges, once a decision is made, it typically requires custom implementation to solve these general problems. Moreover, there is a lack of tool support to aid the custom implementation efforts. We identify this lack of standards and tool support as an opportunity for innovative future work to simplify the evolution of web services.
% section discussion (end)

@@

In this paper, we have summarized our findings of the key challenges with evolutionary web services in four high level categories a) Deployment, b) Implementation, c) Documentation and, d) Communication. For each of these categories, we have compiled a shortlist of challenges in question forms. This list of questions can be treated as a checklist for working on evolving web services. We have also discussed different approaches to solve some of the challenges from our review of the existing literature and industry practices. In our future work, we aim to reconcile and augment these fragmented solutions to provide a standard approach and tool support for evolving web services.
% section conclusion (end)
